Variation in distance between consecutive animal locations

### 2021-03-15 distributions_paper/rudolph_distributions.Rmd
added if statement to easily switch between focus variables

```{r}
library(data)
```

@@

>>>>

```{r}
library(tidyverse)
library(aracari)
library(stringr)
library(fitdistrplus)
library(cowplot)
library(RColorBrewer)
library(gt)

theme_set(theme_bw())
```

```{r}
aracari_df <- readRDS("aracari_df.RDS")
```

```{r eval=FALSE}
# Which individuals will we conside? Need at least a certain number of data points to fit distributions
# Don't remove any yet, just make sure you make it clear how many data points you have for each

aracari_df %>% 
  dplyr::filter(dist != 0) %>% 
  group_by(id) %>% 
  tally()

```

```{r}
# Create my focus variable and only for individuals with 30+ observations

# focus <- "dist"
# focus <- "R2n"
focus <- "mpm"

if(focus == "dist"){
  df <- aracari_df %>%
  dplyr::filter(dist != 0) %>%
  group_by(id) %>%
  filter(., n() >= 30) %>%
  ungroup() %>%
  dplyr::filter(dist != 0) %>% 
    mutate(Bird_ID = paste0("B", id),
           focus = dist)

  dist.used <- c("lnorm", "weibull", "cauchy")
  param <- c("meanlog", "sdlog", "shape", "scale", "location", "scale")
  dist <- c("lnorm", "lnorm", "weibull", "weibull", "cauchy", "cauchy")
}

if(focus == "R2n"){
  df <- aracari_df %>%
  dplyr::filter(R2n != 0) %>%
  group_by(id) %>%
  filter(., n() >= 30) %>%
  ungroup() %>%
  dplyr::filter(R2n != 0) %>% 
    mutate(Bird_ID = paste0("B", id),
           focus = R2n)
}

if(focus == "mpm"){
  df <- aracari_df %>% 
  dplyr::filter(dist != 0) %>% 
  group_by(id) %>%
  filter(., n() >= 30) %>%
  ungroup() %>%
  mutate(tmin = dt/60,
         mpm = dist/tmin,
         Bird_ID = paste0("B", id),
         focus = mpm) %>% 
  dplyr::filter(., mpm != 0)

  dist.used <- c( "exp", "gamma", "weibull", "lnorm")
  param <- c("rate", "shape", "rate", "shape", "scale", "meanlog", "sdlog")
  dist <- c("exp", "gamma", "gamma", "weibull", "weibull", "lnorm", "lnorm")
}

```

@@

### I'll go first with just the distances between locations, might switch to net displacement or net squared displacement later. 
**Visualize the variation and distribution of these distances between locations**

```{r}
 df %>%
  ggplot(., aes((x = focus))) +
  geom_line(aes(x = focus, group = id), stat = "density", alpha = 0.4) +
  geom_line(stat = "density", color = "#1191d1", lwd = 1) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_hline(yintercept = 0, color = "grey") +
  labs(y = "Density", x = paste(focus)) +
  theme_classic()

```

```{r}
df %>%
  ggplot(., aes(x = focus)) +
  facet_wrap(~id) +
  geom_line(stat = "density") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_hline(yintercept = 0, color = "grey") +
  labs(y = "Density", x = paste(focus)) +
  theme_classic()
```

# Fit distribution  
Useful resource to use the package [fitdistrplus](https://cran.r-project.org/web/packages/fitdistrplus/vignettes/FAQ.html)

```{r eval=FALSE}
descdist(df$dist, boot = 1000)
hist(df$dist)

# problem
fitdist(df$dist, distr = "exp")
fitdist(df$dist, distr = "gamma")
#no issue
fitdist(df$dist, distr = "lnorm")
fitdist(df$dist, distr = "weibull")
fitdist(df$dist, distr = "cauchy")

```

```{r}

# Fit the four models for the population level data

pop <- lapply(dist.used, function(x){fitdist(df$focus, distr = x)})

# Get data for qqplot at population level
qqpop <- qqcomp(pop, plotstyle = "ggplot")$data
names(qqpop) <- c("theoretical", "fdist", "empirical")

# Save the population level parameters

build_fits_df <- function(x){

  nm <- deparse(substitute(x))

  x %>%
    map(., `[`, c("estimate", "sd", "loglik", "n")) %>%
    map(as_tibble) %>%
    bind_rows()
}

prms.df <- build_fits_df(pop) %>% 
  mutate(param = param,
         dist = dist,
         data = "pop")

# Repeating the process for data at the individual level
inds <- unique(as.character(df$Bird_ID))

# Fitting the four distributions for each individual and saving the parameters in a dataframe
for(i in 1:length(inds)){
  fit.df <- df %>% 
    dplyr::filter(., Bird_ID == inds[i])

  fit <- lapply(dist.used, function(x){fitdist(fit.df$focus, distr = x)})

  assign(paste(inds[i]), fit)

  fit.prms <- build_fits_df(fit) %>% 
  mutate(param = param,
         dist = dist,
         data = inds[i])

  prms.df <- bind_rows(prms.df, fit.prms)
}

# Getting qqplot data for each individual
inds.models.data <- NULL
ind.models <- list(B1, B13, B19, B22, B28, B29, B3, B30, B49, B5, B7, B84)
# ind.models <- list(B1, B13, B17, B19, B20, B22, B24, B28, B29, B3, B30, B49, B5, B7, B84)
chr.models <- inds
for(i in 1:length(chr.models)){
  x <- ind.models[[i]]

  qqdata <- qqcomp(x, plotstyle = "ggplot")$data %>% 
    mutate(data = chr.models[i])
  names(qqdata) <- c("theoretical", "fdist", "empirical", "individual")
  inds.models.data <- bind_rows(inds.models.data, qqdata)
}
```

```{r ind_QQplots, fig.width=12, fig.height=5}

qqpop %>% 
  mutate(individual = "POPULATION") %>% 
  ggplot(., aes(x = theoretical, y = empirical, color = fdist)) +
  facet_wrap(~individual) +
  #coord_equal() +
  geom_point(shape = 16, alpha = 0.8, size = 4) +
  geom_abline(intercept = 0, slope = 1) +
  labs(x = "Model Quantiles",
         y = "Empirical Quantiles") +
  #scale_color_viridis_d() +
  scale_color_brewer(palette = "Spectral", name = "Distribution") +
  theme(legend.position = c(0.8, 0.2)) +
  NULL -> a

inds.models.data%>% 
  ggplot(., aes(x = theoretical, y = empirical, color = fdist)) +
  #facet_grid(individual~fdist) +
  facet_wrap(~individual)+
    geom_point(shape = 16, alpha = 0.8, size = 3) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Model Quantiles",
         y = "Empirical Quantiles") +
  #scale_color_viridis_d() +
  scale_color_brewer(palette = "Spectral") +
  theme(legend.position = "none") +
  guides(color = guide_legend(title = "Distribution")) +
  #coord_equal() +
  NULL -> b

plot_grid(a,b, rel_widths = c(1,2), labels = "AUTO")

```
